---
title: "Setting up R/Rstudio on DEECA computer"
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# Before You Start

You will need `R` installed on your machine. If your `R` version is <4.0, it is a good time to update. `R` is the programming software that does the computations for your analyses. Using `R` is made much easier by using it through `RStudio` (_I_ntegrated _D_evelopment _E_nvironment). 

If you are using a DEECA machine and and you do not have an open workstation then you may find it difficult to install `R`, `RStudio` and a range of `R` packages. It is recommended you have a workstation setting that allows you to download and install programs (like `R` and `R` packages). This is something you need to apply for through the DEECA IT service portal if you have not already. You should also have `Rtools` installed. 

To install, the order of steps are as below:

1.  Install `R` (\> 4.0.0): <https://cran.rstudio.com/>
    -   install to 'C:\\Windows\\program files' and accept all defaults


2.  Download the version of `Rtools` matching your `R` installation: <https://cran.r-project.org/bin/windows/Rtools/>
    -   install to 'C:\\Windows\\Data' and accept defaults
    
3.  Install `RStudio`: <https://posit.co/download/rstudio-desktop/>
    -   install to 'C:\\Windows\\Data' and accept defaults

4.  Open `Rstudio` as you would any other program to check that it opens.  If so, move onto installing packages below


# Package Dependencies

One of the strengths of `R` is that it is open-source and collaborators can contribute their own code in the form of packages.  For our workshop we will be downloading some packages on the fly, but one `tidyverse` will be across days so please download ahead of time (copy code below and paste into R):

```{r  pkg, eval = FALSE}
install.packages(pkgs = c("tidyverse") )
```


# Workshop prerequisites

# Ummm, pretend that I forgot everything about Rstudio (asking for a mate)...

introduction to Rstudio use...\<\>

<br>

# ...and maybe a refresher on R as well??

introduction to Rstudio use...\<\>

<br>

# Wait, what is this Posit thing? I thought it was Rstudio

switch from Rstudio to Posit...<https://posit.co/>

Rmarkdown to Quarto

Rstudio IDE to Positron <https://github.com/posit-dev/positron/wiki>

<br>

# For f\*sake, now what is this Github thing?

couple notes on navigating Github...<https://github.com/bfanson>

# Now, my brain hurts...

Zen out while checking out the 367 Collins Peregrines....

<iframe width="657" height="370" src="https://www.youtube.com/embed/VLLNNTjaSEA" title="367 Collins Falcons 2024 - South facing" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" referrerpolicy="strict-origin-when-cross-origin" allowfullscreen>

</iframe>

<br>
